---
title: "Linear models"
author: "Richard J. Telford"
date: "May 26, 2016"
output: html_document
---
# Tage Vestlie

## Practical 2



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Modelling Exercises


1. Import the penguin data from the `palmerpenguin` package
```{r ,results = "hide", warning=FALSE}

library(tidyverse)
library(palmerpenguins)
View(penguins)
new_penguins<- penguins%>% 
  drop_na()
```
2. Do big birds have big beaks? Fit a linear model between bill length and body mass, and interpret the coefficients.
```{r}

new_penguins %>% 
  ggplot(aes(x=body_mass_g,y = bill_length_mm))+
  geom_point() +
  stat_smooth(method = "lm" , col = "red") +
  labs(x= "Body mass(g)", y= "bill length(mm)" )

fit_lm1 <- lm(bill_length_mm~body_mass_g, data = new_penguins)
summary(fit_lm1)

```

3. Examine the diagnostics plots. Is everything OK?
```{r}
x11()
par(mfrow=c(2,2))
plot(fit_lm1)
```
<<<<<<< HEAD
3. Calculate the residual sum of squares (find sum the squared difference between the estimates calculated from the coefficients and the observed values).
```{r}
sum(resid(fit_lm1)^2)#shortcut
sum((new_penguins$bill_length_mm-((2.715*10^1)+((4.003*10^-3)*new_penguins$body_mass_g)))^2, na.rm=TRUE)#full calculation
```
=======

3. Calculate the residual sum of squares (find sum the squared difference between the estimates calculated from the coefficients and the observed values).
>>>>>>> 3d0d96dfbbe23dcb977bd5ece5333610e755f88d
4. Recalculate the Residual sum of squares using different values for the coefficients (in the formula yi = Beta + Beta xi + εi).
5. Is the new RSS larger or smaller?
6. Estimate the slope using covariance and variance.
7. Do you get the same result?
8. Add species to the model as a predictor.

```{r} 
new_penguins %>% 
  ggplot(aes(x=body_mass_g,y = bill_length_mm, fill=species))+
  geom_point() +
  stat_smooth(method = "lm" , col = "red") +
  labs(x= "Body mass(g)", y= "bill length(mm)" )
```
9. Does species improve the model?
10. How should the coefficients be interpreted?

