---
title: "Practical 3"
author: "Richard J. Telford"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## GLS

With the built-in `iris` data.

```{r ,results = "hide", warning=FALSE}
library(datasets)
library(tidyverse)
library(nlme)
data(iris)
summary(iris)
```

Make a plot to show how `Petal.Length` varies between species.
```{r}
iris_dt <- iris

iris_dt %>%
  ggplot(aes(x = Species, y = Petal.Length, )) +
  geom_boxplot( colour = c("dark blue","red","pink")) +
  labs(title = "Petal length in different species", x = "Species", y = "Petal lenght(cm)")
```

Find the variance of `Petal.Length` for each species.
```{r}
iris_dt %>% 
  group_by(Species)%>% 
  summarise(var_species = var(Sepal.Length, na.rm =TRUE))
```

Fit an anova using `lm` between `Petal.Length` and species and examine the diagnostic plots.
```{r}
fit_lm1 <- lm(Petal.Length~Species, data = iris_dt)
summary(fit_lm1)

x11()
par(mfrow=c(2,2))
plot(fit_lm1)
```
Fit a `gls` for the same model. Have the coefficients changed? (remember to load ´nlme´ package)
```{r}
fit_gls1 <- gls(Petal.Length~Species, data = iris_dt)
summary(fit_gls1)
```
Fit a `gls` for the same model but allow the variance to be different for each species by adding `varIdent` to the `weights` argument of the model. 
```{r}
fit_gls2 <- gls(Petal.Length~Species, data = iris_dt, weights = varIdent(form = ~ +1|Species))
summary(fit_gls2)
```

Use `AIC` to test if this is a better model.
```{r}
anova(fit_gls1, fit_gls2)
```
