---
title: "Model Selection"
author: "Richard J. Telford"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document: default
  pdf_document: default
---

## hypothesis testing
```{r setup, include=FALSE, warning = FALSE}
library(tidyverse)
library(palmerpenguins)
library(MASS)
library(AICcmodavg)
```
1. Import the palmerpenguin data

```{r}
penguins_df <- penguins %>% 
  drop_na()
```

2. Test the hypothesis that bill length differs between species.
```{r}
penguins_df %>% 
  ggplot(aes(x = species, y = bill_length_mm, fill = sex)) +
  geom_boxplot()



lm_df <- lm(bill_length_mm~species, data = penguins_df)
anova(lm_df)
```
3. Test the hypothesis that bill length differs by sex in addition to species.
```{r}
lm_df2 <- lm(bill_length_mm~species + sex, data = penguins_df)

anova(lm_df,lm_df2)
```
4. How should the p-values be interpreted.

## Exploratory model building 

Normally you would be doing this on a separate data set

5. Use forward selection to find the best model to explain bill length.

6. Build a set of candidate models to explain bill depth using one or two predictors.  

9. Extract the AIC from each  models (hint use function `AIC`). Which is the better model?

10. Calculate the deltaAIC for each model.

11. Calculate the AIC weights for each model. Interpret these weights.


## Collinearity 

12. Make a model predicting bill_length from all other variables. Find the VIF of each predictor. Are there any problem variables? `olsrr::ols_vif_tol`

13. Use `GGally::ggpairs()` to plot the data to try to identify the cause of any high vif.


14. Use `MASS::mvrnorm()` to simulate 100 observation of two predictor variables (x and z) with a given correlation. Simulate a response variable y = b0 + b1x + b2z. Test how the uncertainty in the coefficients changes with the correlation (and hence vif) of the predictor variables.
