---
title: "Bio302 Practical 1 Exploratory Data Analysis"
author: "Richard J. Telford"
date: "May 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1 Peguins

Load the `palmerpenguins` package and see the `penguins` data.

```{r}

library(tidyverse)
library(palmerpenguins)
View(penguins)
```
Find the range, mean, median and variance of variable `bill_length_mm`.
```{r}
new_penguins<- penguins%>% 
  drop_na()

new_penguins %>% 
  summarise(mean = mean(bill_length_mm), 
            medial= median(bill_length_mm),
            range= range(bill_length_mm),
            variance=var(bill_length_mm))

```

What is the standard error of the mean of `bill_length_mm`.
```{r}
new_penguins %>% 
  summarise(sd(bill_length_mm))
```
Plot a histogram of variable `bill_length_mm` + Mark the position of the mean and median.

```{r}
#First create an environment for mean and median. 
bill_length_mean <- new_penguins %>% 
  summarise(mean = mean(bill_length_mm))

bill_length_median <- new_penguins %>% 
  summarise(median = median(bill_length_mm))

#Creating the graph 
new_penguins %>% 
  ggplot(aes(x=bill_length_mm))+
  geom_histogram(binwidth=1, color = "black", fill = "magenta", alpha=0.3) +
  labs(x="Bill length(mm)",y = "Count") +
  geom_vline(xintercept = bill_length_mean$mean, linetype = 2,colour = "maroon", size= 1) +
  geom_vline(xintercept = bill_length_median$median, linetype = 3,colour = "dark blue", size = 1)
```

Plot `bill_length_mm` against `bill_depth_mm`

Find the covariance and correlation of `bill_length_mm` against `bill_depth_mm`.

Is the correlation significant? (hint use cor.test())

How should the p-value be interpreted?

There are three levels in factor `species` Plot `bill_length_mm` against `bill_depth_mm`, coloured by `species`

Is the correlation between `bill_length_mm` against `bill_depth_mm` significant for any of the groups?

This is a sub-group analysis, what are the dangers of this type of analysis?

# Part 2 not penguins

Install the `datasauRus` package with `install.packages("datasauRus")` and load it with `library`.

Dataset `datasaurus_dozen` has `x` and `y` variables and a dataset name. 
For each dataset, calculate the mean and standard deviation of x and y and the correlationn between x and y.